### 2015-04-12 src/intro/intro.tex
Fixed figure references

\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/intro/figures/vestibular.pdf}
    \caption{Schematic view of the inner ear complex consisting of the auditory system (cochlea) and the vestibular system. The vestibular system has three angular velocity (semicircular canals) and two linear acceleration (saccule and utricle) sensors.}
    \label{intro:fig6}
\end{figure}

@@

The semicircular canals measure the three-dimensional rotational velocity of the head. Each side of the head contains three orthogonally oriented canals allowing for rotation to be perceived in all three dimensions. Each of the six canals consists of a circular tube filled with a fluid known as endolymph (see \figref{intro:fig3}A). One part of the tube, the ampulla, is a bit thicker than the rest and contains a membrane, the cupula, that separates the fluid. When the head rotates, the fluid stays behind because of its inertia which in turn causes the membrane to deflect (see \figref{intro:fig3}B). While the inertial fluid motion suggests that the cupula should be sensitive to angular acceleration, reactive forces resulting from the fluid motion, such as endolymph viscosity and cupular elasticity,  cause the cupular deflection to reflect angular velocity instead \cite{goldberg2012}.

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/intro/figures/canals.pdf}
    \caption{Detailed view of a semicircular canal \panelref{A} in rest and \panelref{B} during rotation. Inertia of the endolymph in the canal causes the cupula to deflect during rotation.}
    \label{intro:fig3}
\end{figure}

@@

\subsubsection{Otoliths}
The linear acceleration and gravitational forces are measured by the two otolith organs on either side of the head: the saccule and the utricule. Each otolith consists of an endolymph filled compartment containing calcium carbonate crystals known as otoconia. Due to their high inertia, these otoconia fall behind during linear acceleration, and move ahead during deceleration. In addition, they fall downward as a result of the gravitational field. The otoconia are mounted on top of a flexible polysaccharide gel, in which hair cells, similar to those found in the semicircular canals, are partly embedded \cite{goldberg2012}.

@@

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/intro/figures/otoliths.pdf}
    \caption{Detailed view of an otolith \panelref{A} in rest, \panelref{B} during head tilt, and \panelref{C} during linear acceleration. The high inertia of the otoconia keeps them from moving rapidly during both head tilt as well as linear acceleration. As the effect is physically identical in either case, head tilt and linear acceleration cannot be disambiguated based on the otolith signal alone.}
    \label{intro:fig7}
\end{figure}

@@

\subsubsection{Vection}
When we move through the environment, the image of the world on our retina shifts. This large-field shift pattern, also known as optic flow, depends on the movement being made. Lateral translation (\figref{intro:fig2}A) for example causes a different pattern than roll rotation (\figref{intro:fig2}B). At the turn of the 19th century von Helmholtz \citeyear{vonhelmholtz1867} recognized the importance of these flow signals for self-motion perception. In some cases the optic flow signal is so strong that it causes a percept of self-motion in stationary participants, called vection \cite{dichgans1978}. This is experienced, for example, when sitting on the train and a neighbouring train starts to move. This effect is much less likely to occur when e.g. on the platform, suggesting that the vection signal is integrated with prior knowledge about the environment before causing self-motion perception \cite{andersen1985,lepecq1995}

@@

    \caption{Vection pattern during \panelref{A} lateral translation and \panelref{B} rotation. During translation nearby targets (red) have a larger retinal displacement compared to far away ones (blue). Retinal displacement across rotation does not depend on target distance. }
    \label{intro:fig2}
\end{figure}

@@

\subsubsection{Landmarks}
In addition to relative cues, the brain seems to use absolute cues in both the perception of gravity and self-motion. For the perception of gravity, it makes use of the fact that many lines within the world are aligned with either the horizon or gravity. Straight lines therefore acts as a priors attracting the perceived direction of gravity towards them. A special case is the rod-and-frame illusion (\figref{intro:fig9}), where the perceived angle of a rod relative to gravity is affected by the orientation of the frame that contains it \cite{witkin1948}.

\begin{figure}
	\includegraphics[width=1.0\textwidth]{src/intro/figures/treeframe.pdf}
	\caption{\panelref{A} Trees are aligned with gravity, making them ideal cues for orientation. \panelref{B} The rod-and-frame illusion; even though the frame is rotated, participants might perceive the line as being slanted.}
	\label{intro:fig9}
\end{figure}

Similarly, our position within the world can be established using world-fixed landmarks. With the exception of animals and vehicles, most items in the world rarely change position. By comparing our current visual scene (e.g. Sequoia trees in \figref{intro:fig1}) with our knowledge about the environment we can determine our location (in this case northern California). Because the experiments outlined in this thesis were performed in near darkness, these absolute navigational cues can be considered less relevant.

@@

In 1963, \citeauthor{guedry1963} reported a substantial underestimation of displacement when their observers watched a small body-fixed target compared to displacements in the dark. Because of the VOR, eye movement in darkness are larger than those in the body-fixed condition. The underestimation in the body-fixed condition can therefore be interpreted as the involvement of eye movements, although the authors did not suggest this 
Studies on postural sway also indicate a role of eye movement is self-motion perception. Making eye movements causes postural sway to increase, suggesting that eye movements influence self-motion perception \cite{glasauer2005,rogrigues2015}.

%\begin{figure}
%	\includegraphics[width=1.0\textwidth]{src/intro/figures/...}
%\end{figure}

One complication is that the magnitude, $e$, of these eye movements not only depend on the translation amplitude, $T$, but also on the fixation depth, $d$. When fixating position $(x, d)$, the eye position at starting position is $e_0 = atan⁡(d/x) \approx d/x$. By subtracting the eye position after the translation, $e_1 = atan⁡(d/(x+T)) \approx d/(x+T)$, we obtain the eye movement amplitude, $e \approx d/T$. In Chapter 4 we explore whether the brain takes this geometry into account and compensates for fixation depth when using eye movements in self-motion perception.

@@

The vestibulo-ocular reflex (VOR) is driven by the short-latency vestibular signal and compensates for both rotation and translation of the head. The rotational and translational part of the VOR have their own dedicated reflex arcs. The translational VOR, or TVOR, is driven by the otolith signal while the rotational VOR, or RVOR, is mainly driven by the semicircular canals. The RVOR operates during head rotation by counter rotating the eyes in the opposite direction to the head. Because its gain is about one, eye velocity is about equal, but of opposite sign, to the head velocity. The TVOR operates during head translation, when the head moves orthogonally to the line of sight. In contrast to the RVOR, the TVOR needs to take fixation depth into account. Simple geometry shows that the ideal TVOR response is inversely proportional to fixation distance, $e_vor = atan⁡(d/T)$. As a result, no compensatory eye movements are required when fixating targets at infinity to maintain a stable retinal image.

@@

    \caption{Subjective visual vertical task (SVV); \panelref{A} Graphical representation of the SVV task. \panelref{B} Bias and \panelref{C} uncertainty as a function of roll angle in a typical participant (circles). The dashed line represents the typical pattern.}
    \label{intro:fig4}
\end{figure}

>>>>

\begin{figure}
    \includegraphics[width=0.5\textwidth]{src/intro/figures/vestibular.pdf}
    \caption{Schematic view of the inner ear complex consisting of the auditory system (cochlea) and the vestibular system. The vestibular system has three angular velocity (semicircular canals) and two linear acceleration (saccule and utricle) sensors.}
    \label{intro:vestibular}
\end{figure}

@@

The semicircular canals measure the three-dimensional rotational velocity of the head. Each side of the head contains three orthogonally oriented canals allowing for rotation to be perceived in all three dimensions. Each of the six canals consists of a circular tube filled with a fluid known as endolymph (see \figref{intro:fig:canals}A). One part of the tube, the ampulla, is a bit thicker than the rest and contains a membrane, the cupula, that separates the fluid. When the head rotates, the fluid stays behind because of its inertia which in turn causes the membrane to deflect (see \figref{intro:fig:canals}B). While the inertial fluid motion suggests that the cupula should be sensitive to angular acceleration, reactive forces resulting from the fluid motion, such as endolymph viscosity and cupular elasticity,  cause the cupular deflection to reflect angular velocity instead \cite{goldberg2012}.

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/intro/figures/canals.pdf}
    \caption{Detailed view of a semicircular canal \panelref{A} in rest and \panelref{B} during rotation. Inertia of the endolymph in the canal causes the cupula to deflect during rotation.}
    \label{intro:fig:canals}
\end{figure}

@@

\subsubsection{Otoliths}
The linear acceleration and gravitational forces are measured by the two otolith organs on either side of the head: the saccule and the utricule. Each otolith consists of an endolymph filled compartment containing calcium carbonate crystals known as otoconia (\figref{intro:fig:otoliths}A). Due to their high inertia, these otoconia fall behind during linear acceleration (\figref{intro:fig:otoliths}B), and move ahead during deceleration. In addition, they fall downward as a result of the gravitational field (\figref{intro:fig:otoliths}C). The otoconia are mounted on top of a flexible polysaccharide gel, in which hair cells, similar to those found in the semicircular canals, are partly embedded \cite{goldberg2012}.

@@

\begin{figure}
    \includegraphics[width=1.0\textwidth]{src/intro/figures/otoliths.pdf}
    \caption{Detailed view of an otolith \panelref{A} in rest, \panelref{B} during head tilt, and \panelref{C} during linear acceleration. The high inertia of the otoconia keeps them from moving rapidly during both head tilt as well as linear acceleration. As the effect is physically identical in either case, head tilt and linear acceleration cannot be disambiguated based on the otolith signal alone.}
    \label{intro:fig:otoliths}
\end{figure}

@@

\subsubsection{Vection}
When we move through the environment, the image of the world on our retina shifts. This large-field shift pattern, also known as optic flow, depends on the movement being made. Lateral translation (\figref{intro:fig:vection}A) for example causes a different pattern than roll rotation (\figref{intro:fig:vection}B). At the turn of the 19th century von Helmholtz \citeyear{vonhelmholtz1867} recognized the importance of these flow signals for self-motion perception. In some cases the optic flow signal is so strong that it causes a percept of self-motion in stationary participants, called vection \cite{dichgans1978}. This is experienced, for example, when sitting on the train and a neighbouring train starts to move. This effect is much less likely to occur when e.g. on the platform, suggesting that the vection signal is integrated with prior knowledge about the environment before causing self-motion perception \cite{andersen1985,lepecq1995}

@@

    \caption{Vection pattern during \panelref{A} lateral translation and \panelref{B} rotation. During translation nearby targets (red) have a larger retinal displacement compared to far away ones (blue). Retinal displacement across rotation does not depend on target distance. }
    \label{intro:fig:vection}
\end{figure}

@@

\subsubsection{Landmarks}
In addition to relative cues, the brain seems to use absolute cues in both the perception of gravity and self-motion. For the perception of gravity, it makes use of the fact that many lines within the world are aligned with either the horizon or gravity (\figref{intro:fig:treeframe}A). Straight lines therefore acts as a priors attracting the perceived direction of gravity towards them. A special case is the rod-and-frame illusion (\figref{intro:fig:treeframe}B), where the perceived angle of a rod relative to gravity is affected by the orientation of the frame that contains it \cite{witkin1948}.

\begin{figure}
	\includegraphics[width=1.0\textwidth]{src/intro/figures/treeframe.pdf}
	\caption{\panelref{A} Trees are aligned with gravity, making them ideal cues for orientation. \panelref{B} The rod-and-frame illusion; even though the frame is rotated, participants might perceive the line as being slanted.}
	\label{intro:fig:treeframe}
\end{figure}

Similarly, our position within the world can be established using world-fixed landmarks. With the exception of animals and vehicles, most items in the world rarely change position. By comparing our current visual scene (e.g. Sequoia trees in \figref{intro:fig:treeframe}A) with our knowledge about the environment we can determine our location (in this case northern California). Because the experiments outlined in this thesis were performed in near darkness, these absolute navigational cues can be considered less relevant.

@@

In 1963, \citeauthor{guedry1963} reported a substantial underestimation of displacement when their observers watched a small body-fixed target compared to displacements in the dark. Because of the VOR, eye movement in darkness are larger than those in the body-fixed condition. The underestimation in the body-fixed condition can therefore be interpreted as the involvement of eye movements, although the authors did not suggest this 
Studies on postural sway also indicate a role of eye movement is self-motion perception. Making eye movements causes postural sway to increase, suggesting that eye movements influence self-motion perception \cite{glasauer2005,rodrigues2015}.

One complication is that the magnitude, $e$, of these eye movements not only depend on the translation amplitude, $T$, but also on the fixation depth, $d$. When fixating position $(x, d)$, the eye position at starting position is $e_0 = atan(d/x) \approx d/x$. By subtracting the eye position after the translation, $e_1 = atan(d/(x+T)) \approx d/(x+T)$, we obtain the eye movement amplitude, $e \approx d/T$. In Chapter 4 we explore whether the brain takes this geometry into account and compensates for fixation depth when using eye movements in self-motion perception.

@@

The vestibulo-ocular reflex (VOR) is driven by the short-latency vestibular signal and compensates for both rotation and translation of the head. The rotational and translational part of the VOR have their own dedicated reflex arcs. The translational VOR, or TVOR, is driven by the otolith signal while the rotational VOR, or RVOR, is mainly driven by the semicircular canals. The RVOR operates during head rotation by counter rotating the eyes in the opposite direction to the head. Because its gain is about one, eye velocity is about equal, but of opposite sign, to the head velocity. The TVOR operates during head translation, when the head moves orthogonally to the line of sight. In contrast to the RVOR, the TVOR needs to take fixation depth into account. Simple geometry shows that the ideal TVOR response is inversely proportional to fixation distance, $e_{vor} = atan⁡(d/T)$. As a result, no compensatory eye movements are required when fixating targets at infinity to maintain a stable retinal image.

@@

    \caption{Subjective visual vertical task (SVV); \panelref{A} Graphical representation of the SVV task. \panelref{B} Bias and \panelref{C} uncertainty as a function of roll angle in a typical participant (circles). The dashed line represents the typical pattern.}
    \label{intro:fig5}
\end{figure}
